Testing new Pauly Cessational Growth Function

### 2016-04-13 manuscript_v0.Rmd
Initial work on manuscript

---
title: "Testing new Pauly Cessational Growth Function"
author: "Derek H. Ogle"
date: "April 9, 2016"
output: 
  pdf_document: 
    fig_caption: yes
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FSA)
library(FSAsim)

# setup figure, table, and equation captioning
library(captioner)
figcaps <- captioner(prefix="Figure")
tblcaps <- captioner(prefix="Table")
eqncaps <- captioner(prefix="Equation")
vbSO <- vbFuns("Somers2")
```

@@

The growth of many animals exhibits seasonal oscillations in response to seasonal changes in temperature, light, and food supplies ().  Several functions have been proposed to model seasonal growth.  The most popular of these functions, from Somers (1988) but carefully reiterated in GBO (), is

\[ L_{t} = L_{\infty}(1-e^{-K(t-t_{0})}-S(t)+S(t_{0}))  \]

with $S(t) = \frac{CK}{2\pi}sin(2\pi(t-t_{s}))$ so that $S(t_{0}) = \frac{CK}{2\pi}sin(2\pi(t_{0}-t_{s}))$ and where $L(t)$ is the expected or average length at time (or age) $t$, $L_{\infty}$ is the model asymptote for average length, $K$ is a measure of the exponential rate of approach to the asymptotic length (Schnute and Fournier 1980), $t_{0}$ is the theoretical time or age (generally negative) at which the average length would be zero, $C$ modulates the amplitude of the growth oscillations and corresponds to the proportion of decrease in growth at the depth of the oscillation (i.e., "winter"), and $t_{s}$ is the time between time 0 and the start of the convex portion of the first sinusoidal growth oscillation (i.e., the inflection point).  If $C$=0, then there is no seasonal oscillation and this function reduces to the typical von Bertalanffy growth function.  If $C$=1, then growth completely stops once a year at the "winter-point" ($WP=t_{s}+0.5$), whereas values of 0&gt;$C$&lt;1 result in reduced, but not stopped, growth during the winter.

Values of C[1 (or\0)
allow seasonal decreases in average length-at-age,
and might not seem realistic in organisms whose
skeletons largely preclude shrinkage (Pauly et al.
1992), but could result from size-dependent overwinter
mortality.

1. Seasonal Growth
1. Somer's Seasonal Growth Function (SoSVBGF)

```{r echo=FALSE, fig.width=5, fig.height=4, fig.cap="Example growth trajectory of Equation X using Linf=30, Kpr=0.7, t0=-0.05, WP=0.8, and NGT=0.4.  The start of each no growth time (i.e, Annual Age+WP) is marked by a red vertical line and the end of each no growth time (i.e., Annual Age+WP+NGT)."}
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
WP <- 0.8; C <- 1
Linf <- 30; K <- 0.7; t0 <- -0.05
Scf <- c(Linf,K,t0,C,WP)
t <- seq(0,5,length.out=299)
SL <- vbSO(t,Linf=Scf)
plot(SL~t,type="l",lwd=2,ylim=c(0,30),ylab="Length",xlab="Age (years)")
Scf <- c(Linf,K,t0,C=2,WP)
SL <- vbSO(t,Linf=Scf)
lines(SL~t,lwd=2,lty=2)
Scf <- c(Linf,K,t0,C=0.2,WP)
SL <- vbSO(t,Linf=Scf)
lines(SL~t,lwd=2,lty=3)

WPs <- 0:5+WP
LatWPs <- vbSO(WPs,Linf=Scf)
segments(WPs,0,WPs,LatWPs,lty=3,lwd=2,col="red")

legend("topleft",c("C=1","C=2","C=0.2"),lwd=2,lty=1:3,bty="n",inset=-0.02)
```

1. Pauly's Seasonal Growth Function (PaSVBGF)
    1. Define terms
    1. Explain t-prime
    1. Explain K-prime

```{r echo=FALSE, fig.width=5, fig.height=4, fig.cap="Example growth trajectory of Equation X using Linf=30, Kpr=0.7, t0=-0.05, WP=0.8, and NGT=0.4.  The start of each no growth time (i.e, Annual Age+WP) is marked by a red vertical line and the end of each no growth time (i.e., Annual Age+WP+NGT)."}
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
WP <- 0.8; NGT <- 0.4
Linf <- 30; Kpr <- 0.7; t0 <- -0.05
Pcf <- c(Linf,Kpr,t0,WP,NGT)
t <- seq(0,5,length.out=299)
PL <- paulySC(t,Linf=Pcf)
plot(PL~t,type="l",lwd=2,ylim=c(0,30),ylab="Length",xlab="Age (years)")
WPs <- 0:5+WP
LatWPs <- paulySC(WPs,Linf=Pcf)
segments(WPs,0,WPs,LatWPs,lty=3,lwd=2,col="red")
segments(WPs+NGT,0,WPs+NGT,LatWPs,lty=3,lwd=2,col="blue")
legend("topleft",c("WP=0.8","NGT=0.4"),bty="n",inset=-0.02,text.col=c("red","blue"))
```

>>>>

---
title: "Testing new Pauly Cessational Growth Function"
author: "Derek H. Ogle"
date: "April 9, 2016"
output:
  html_document: 
    fig_height: 3.5
    fig_width: 4
  pdf_document:
    fig_caption: yes
csl: american-fisheries-society.csl
bibliography: SeasonalGrowth.bib
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(34738423)  # for reproducibility
library(FSA)
library(FSAsim)

# setup figure, table, and equation captioning
library(captioner)
figcaps <- captioner(prefix="Figure")
tblcaps <- captioner(prefix="Table")
eqncaps <- captioner(prefix="Equation")
eqncaps("VBMain")
eqncaps("SomersMod")
eqncaps("PaulyMod")
eqncaps("PaulyMod2")

vbSO <- vbFuns("Somers2")
```

@@

The growth of many animals exhibits seasonal oscillations in response to seasonal changes in temperature, light, and food supplies ().  Several functions have been proposed to model seasonal growth.  The most popular of these functions, from XXX and @Somers1988 and carefully reiterated in @GarciaBerthouetal2012, is

$$ L_{t} = L_{\infty}(1-e^{-q}) \quad \quad \text{`r paste0("(",eqncaps("VBMain",display="num"),")")`} $$

with

$$ q=K(t-t_{0})+\frac{CK}{2\pi}sin(2\pi(t-t_{s}))-\frac{CK}{2\pi}sin(2\pi(t_{0}-t_{s})) \quad \quad \text{`r paste0("(",eqncaps("SomersMod",display="num"),")")`} $$

where $L(t)$ is the expected or average length at time (or age) $t$, $L_{\infty}$ is the model asymptote for average length, $K$ is a measure of the exponential rate of approach to the asymptotic length [@SchnuteFournier1980], $t_{0}$ is the theoretical time or age (generally negative) at which the average length would be zero, $C$ modulates the amplitude of the growth oscillations and corresponds to the proportion of decrease in growth at the depth of the oscillation (i.e., "winter"), and $t_{s}$ is the time between time 0 and the start of the convex portion of the first sinusoidal growth oscillation (i.e., the inflection point).  If $C$=0, then there is no seasonal oscillation and this function reduces to the typical von Bertalanffy growth function (`r figcaps("SomersExPlot",display="cite")`).  If $C$=1, then growth completely stops once a year at the "winter-point" ($WP=t_{s}+0.5$), whereas values of 0&gt;$C$&lt;1 result in reduced, but not stopped, growth during the winter (`r figcaps("SomersExPlot",display="cite")`).

\

```{r echo=FALSE, results="hide"}
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
WP <- 0.8; Linf <- 30; K <- 0.3; t0 <- -0.05
C <- c(0,0.5,1,2)
clr <- col2rgbt(rep("black",4),1/(1:4))
t <- seq(0,3,length.out=299)
plot(vbSO(t,Linf=c(Linf,K,t0,C=C[1],WP))~t,type="l",lwd=2,col=clr[1],
     ylim=c(0,20),ylab="Length",xlab="Age (years)",xaxt="n")
axis(1,0:3)
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[2],WP))~t,lwd=2,col=clr[2])
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[3],WP))~t,lwd=2,col=clr[3])
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[4],WP))~t,lwd=2,col=clr[4])

#WPs <- 0:5+WP
#LatWPs <- vbSO(WPs,Linf=Scf)
#segments(WPs,0,WPs,LatWPs,lty=3,lwd=2,col="red")

legend("topleft",paste("C",C,sep="="),lwd=2,col=clr,bty="n",inset=-0.02)

figcaps("SomersExPlot",paste0("Example growth trajectory of ",eqncaps('VBMain',display='cite')," using ",eqncaps('SomersMod',display='cite')," with $L_{\\infty}$=",Linf,", $K$=",K,", $t_{0}$=",t0,", $t_{s}$=",WP-0.5," (or $WP$=",WP,"), and four different values of $C$."))
```

`r figcaps("SomersExPlot")`

\

Values of $C$&gt;1 (or &lt;0) allow seasonal decreases in average length-at-age (`r figcaps("SomersExPlot",display="cite")`).  A seasonal decrease in average length-at-age could result from size-dependent overwinter mortality [@GarciaBerthouetal2012], but this growth trajectory is unlikely for organisms whose skeletons largely preclude shrinkage [@Paulyetal1992].

@Paulyetal1992 introduced a modification of `r eqncaps("SomersMod")` that modeled a true no-growth period and that smoothly transitioned into and out of the no-growth period.  Specifically, their function is `r eqncaps("VBMain")` with

\

$$ q=K(t^{'}-t_{0})+\frac{K(1-NGT)}{2\pi}sin(\frac{2\pi}{1-NGT}(t^{'}-t_{s}))-\frac{K(1-NGT)}{2\pi}sin(\frac{2\pi}{1-NGT}(t_{0}-t_{s})) \quad \quad \text{`r paste0("(",eqncaps("PaulyMod",display="num"),")")`} $$

\

\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (in fractions of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$."  @Paulyetal1992 introduced $K^{'}$ to `r eqncaps("PaulyMod",display="cite")` as the units of $K$ change from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")` because $t^{'}$&lt;$t$ for any $t$.  We further note here that $K^{'}$&gt;$K$ because annual growth is compressed into a shorter period ($1-NGT$) in `r eqncaps("PaulyMod",display="cite")` when $NGT$&gt;0.

@Paulyetal1992 derived  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at one point), replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the growth period), and replacing $K$ with $K^{'}$.  Furthermore, their modification can be described geometrically (though not algorithmically) in two steps.  First, the seasonal growth function combined in Equations `r eqncaps("VBMain",display="num")` and `r eqncaps("SomersMod",display="num")` with $C$=1 is fit to the observed lengths and ages that have had the cumulative no-growth time subtracted (i.e., the $t^{'}$).  The growth trajectory is then separated at the $WP$ and a segment with a slope of zero that is $NGT$ units long is inserted into the trajectory.  This forms a trajectory that smoothly transitions into and out of a no-growth period (`r figcaps("PaulyExPlot")).

\

```{r echo=FALSE, results="hide"}
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
clrs <- c("black",col2rgbt("black",1/2))
WP <- 0.8; NGT <- 0.3
Linf <- 30; Kpr <- 0.4; t0 <- -0.05
Pcf <- c(Linf,Kpr,t0,WP,NGT)
t <- seq(0,3,length.out=299)
PL <- paulySC(t,Linf=Pcf)
plot(PL~t,type="l",lwd=2,ylim=c(0,20),ylab="Length",xlab="Age (years)")
WPs <- 0:3+WP
LatWPs <- paulySC(WPs,Linf=Pcf)
segments(WPs,LatWPs-2,WPs,LatWPs+2,lty=3,lwd=2,col=clrs[1])
segments(WPs-NGT/2,LatWPs-1,WPs-NGT/2,LatWPs+1,lty=3,lwd=2,col=clrs[2])
segments(WPs+NGT/2,LatWPs-1,WPs+NGT/2,LatWPs+1,lty=3,lwd=2,col=clrs[2])
figcaps("PaulyExPlot",paste0("Example growth trajectory of ",eqncaps('VBMain',display='cite')," using ",eqncaps('PaulyMod',display='cite')," with $L_{\\infty}$=",Linf,", $K^{'}$=",Kpr,", $t_{0}$=",t0,", $WP$=",WP,", and $NGT$=",NGT,".  Each $WP$ is marked by the longer vertical lines and the beginning and ending of the no-growth times are marked by the paired shorter vertical lines."))
```

`r figcaps("PaulyExPlot")`

\
